---
title: 'Differential methylation: Maltreatment effects'
author: "Vera N. Karlbauer"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/Users/vera_karlbauer/Documents/sib-best-practices-course')
```

# Code extract for SIB "Best Practices in Programming" course

## What this script does:
- Analyzes effects of maltreatment on 49 CpG sites via linear regression. 
- Runs linear model for first timepoint (T0).  
- Exports results as .txt table.  
- The functions looping the linear model over 49 CpGs, specifying all relevant covariates, and providing a dataframe with adjusted R squared, regression coefficients, p-values, and fdr-corrected p-values are specified in the script 02_stats/00_functions.R .  
- All analyses are corrected for age, sex, genetic components (C1, C2, C3), cell type counts (CD34, CD14, buccal), prenatal smoke exposure, and prenatal alcohol exposure. 

## Setup

### Step 0: Load / install packages

```{r}
# install packages (commented out)
# install.packages("rmarkdown")
# install.packages("dplyr")
# install.packages("lme4")
# install.packages("stringr")

library(rmarkdown)
library(dplyr)
library(lme4)
library(stringr)

```

### Step 1: Source functions

```{r}
source("./02_analysis/00_functions.R")

```


### Step 2: Load data

```{r}
# all_t0 = methylation & covariates for T0
load("./01_data/all_t0.rda")

```

# Analysis

## Maltreatment effects at T0

### Run statistical analysis

```{r}
# loop linear model with pre-specified covariates and maltreatment as main predictor over 49 CpGs
loop_lm_simple(data = all_t0, predictor_name = "Case.Control", object_name = "diffmeth_malt_t0")

# count number of uncorrected significant hits
count(diffmeth_malt_t0, sig_uncorrected)

# count number of fdr-corrected significant hits
count(diffmeth_malt_t0, sig_corrected)

```

### Export results

```{r}
# round values
diffmeth_malt_t0 <- diffmeth_malt_t0 %>%
  mutate(across(c("adj.R", "p", "coef", "p_adjust"), round, 3))

# export
write.table(diffmeth_malt_t0, "./03_results/results_diffmeth_malt_t0.txt", row.names = FALSE)

```

